---
title: "Homework 5"
author: "Kodiak Soled"
date: "11/2/2019"
output: github_document
always_allow_html: yes
---

```{r setting up document, include = FALSE}
library(tidyverse)
library(patchwork)
library(viridis)
library(kableExtra)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d

scale_fill_discrete = scale_fill_viridis_d

theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

## Problem 1

#### Read in Data

First, I read in the missing data. 

```{r loading in missing dataset}
set.seed(10)

iris_with_missing = 
  iris %>% 
  map_df(~replace(.x, sample(1:150, 20), NA)) %>%
  mutate(Species = as.character(Species))
```

#### Create a Function to Address Missing Data

In order to fill in the missing data, I first needed to write a `function` that:

* Took a vector as an argument
* `Replace`d missing values:
    * of numeric variables with the mean of non-missing values
    * of character variables with `"virginica"`
* Returned the resulting vector

```{r write a function to replace missing values}
tidy_iris = function(x) {
  
  if (!is.numeric(x) & !is.character(x)) {
    stop("Argument x should be numeric or character")
  } else if (is.character(x)) {
    x = replace(x, is.na(x) == TRUE, "virginica")
  } else if (is.numeric(x)) {
    mean_x = mean(x, na.rm = TRUE)
    x = replace(x, is.na(x) == TRUE, mean_x)
  }
  
}
```

#### Create a Map Statement

Then, I applied this function to the `iris_with_missing` dataset using a `map` statement and `bind_rows` in order to combine the 5 datasets into a single one:

```{r write a map statement, warning = FALSE}
output = 
  map(iris_with_missing, tidy_iris) %>%
  bind_rows()
```

#### Table of New Iris Data

Here is a reader-friendly table the first 30 observations of the replaced iris dataset:

```{r resulting table}
head(output, n = 30) %>%
  janitor::clean_names() %>%
  knitr::kable(digits = 1, caption = "Iris Dataset (First 30 Observations)") %>%
  kable_styling(bootstrap_options = c("striped", "condensed", font_size = 12))
```

## Problem 2

#### Load in Data files

In order to create a tidy dataframe containing data from all participants, including the subject ID, arm, and observations over time, I first loaded in the 20 seperate datafiles using `list.files` (created a dataframe with the file names):

```{r}
study_data_2 = 
  list.files(path = "./data_hw5", full.names = FALSE) %>%
  tibble::enframe(name = NULL)
study_data_2
```

#### Create Tidy Dataframe

Then, I iterated over the file names and read in the data (`read_csv`) for each of the 20 subjects (from 20 strings to 1 string using `str_c`) using `purrr::map` (and saving the result as a new variable in the dataframe). I then `unnest`ed the variable "data" so I could see the measurements for each subject across the eight weeks. I then tidied the dataset so that there were `seperate` variables for study group (control vs. experimental) and subject id and removed the ".csv" in the name using `mutate` and `str_replace`. Finally, I put the dataframe into a reader-friendly table using `knitr::kable` and made sure all the measurements were rounded to 2 decimal places using `digits = 2`.

```{r write using map, message = FALSE}
df_study_data_2 = 
  study_data_2 %>%
  mutate(
    data = map(value, ~read_csv(str_c("./data_hw5/", .x)))
  ) %>%
  unnest(cols = data) %>%
  separate(value, into = c("arm", "subject_id"), sep = "_") %>%
  mutate(subject_id = str_replace(subject_id, ".csv", ""))
df_study_data_2 %>%
  knitr::kable(digits = 2, caption = "Tidy Study Dataframe") %>%
  kable_styling(bootstrap_options = c("striped", "condensed", font_size = 12))

```