---
title: "Homework 5"
author: "Kodiak Soled"
date: "11/2/2019"
output: github_document
always_allow_html: yes
---

```{r setting up document, include = FALSE}
library(tidyverse)
library(patchwork)
library(viridis)
library(kableExtra)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d

scale_fill_discrete = scale_fill_viridis_d

theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

## Problem 1

#### Read in Data

First, I read in the missing data. 

```{r loading in missing dataset}
set.seed(10)

iris_with_missing = 
  iris %>% 
  map_df(~replace(.x, sample(1:150, 20), NA)) %>%
  mutate(Species = as.character(Species))
```

#### Create a Function to Address Missing Data

In order to fill in the missing data, I first needed to write a `function` that:

* Took a vector as an argument
* `Replace`d missing values:
    * of numeric variables with the mean of non-missing values
    * of character variables with `"virginica"`
* Returned the resulting vector

```{r write a function to replace missing values}
tidy_iris = function(x) {
  
  if (!is.numeric(x) & !is.character(x)) {
    stop("Argument x should be numeric or character")
  } else if (is.character(x)) {
    x = replace(x, is.na(x) == TRUE, "virginica")
  } else if (is.numeric(x)) {
    mean_x = mean(x, na.rm = TRUE)
    x = replace(x, is.na(x) == TRUE, mean_x)
  }
  
}
```

#### Create a Map Statement

Then, I applied this function to the `iris_with_missing` dataset using a `map` statement and `bind_rows` in order to combine the 5 datasets into a single one:

```{r write a map statement, warning = FALSE}
output = 
  map(iris_with_missing, tidy_iris) %>%
  bind_rows()
```